---
title: "Introduction to R"
author: 
date: "11/11/2021"
output: pdf_document 
---

# Day 1
- Installation of R and RStudio
- Overview on the R and RStudio layout and menus
- Installation of packages required for the training


# Day 2
- Recap
- Data types and data structures in R
- Setting of working directory
- Reading data into R 
- Helper functions for exploring data in R
- Writing data out of R

# Day 3
- Recap
- Data Cleaning in R

# Day 4
- Recap
- Data Manipulation in R 

# Day 5
- Recap
- Data Visualization in R


# Duration for the training
- Two hours per day

```{r}
ggplot(dat1,aes(name,value,fill=LoS))+stat_summary(fun = 'mean',position = 'dodge',geom = 'bar')+stat_summary(geom='errorbar',fun.data = 'mean_se',width=.2,position = position_dodge(.9))
```

```{r}
ggplot(dat1,aes(name,value,fill=dose))+
    stat_summary(geom='errorbar',fun.data = 'mean_se',width=1,position = position_dodge(.9))+
    stat_summary(fun = 'mean',position = 'dodge',geom = 'bar')+scale_fill_aaas()+scale_y_continuous(expand = c(0,0))+facet_grid(~LoS)+coord_flip()
```

